% -*- mode: LaTeX; TeX-PDF-mode: t; -*- # Tell emacs the file type (for syntax coloring)
\newcommand{\econtexRoot}{..}\input{\econtexRoot/Resources/.econtexPaths.tex}
\documentclass{\classes/econtex}
\usepackage{graphicx}
\usepackage{\LaTeXInputs/local}
\usepackage{\packages/econtexSetup}
\usepackage{\packages/pdfsuppressruntime}

% Allow links to items inside itemize environment
% https://tex.stackexchange.com/questions/81949/can-individual-items-in-an-itemize-list-be-labelled-and-hyperref-linked/81952#81952
% \AtBeginEnvironment{itemize}{\apptocmd{\item}{\phantomsection}{}{\errmessage{couldn't patch item}}}
\renewcommand{\thepage}{} % Get rid of page numbers, which don't convert to md or html

\providecommand\phantomsection{}
\provideboolean{MyNotes}\setboolean{MyNotes}{false} % Whether to show marginalia

\pagestyle{plain}
\hypersetup{pageanchor=false} % prevents useless warning messages
\begin{document}
\hfill{\tiny \jobname, \today} \vspace{.1in}

\begin{verbatimwrite}{\jobname.title}
  Frequently Asked Questions About the Job Market
\end{verbatimwrite}

\section*{\Large Frequently Asked Questions About the Job Market}\medskip\medskip

\centerline{\today}\medskip\medskip

% \ifthenelse{\boolean{MyNotes}}{\marginpar{\tiny }}{}
\begin{enumerate}

\item Who is in charge of the job market process?
  \begin{quote}
    You are.  You are responsible for your own job market search.  (The economics PhD job market is a very capitalist institution).

    The~\href{\Notationpage}{JMPO}~is the overall coordinator of the process, responsible (for example) for updating these notes.  But you should \textit{not} wait for specific reminders and prodding by the JMPO before doing the things you are supposed to do. If you fail to do things that are in your own best interest, and which nobody will do for you, there is nobody to blame but yourself.

  \end{quote}

\item  Who is eligible to go on the market?
  \begin{quote}
    
    Students who

    \begin{enumerate}
      \item have the approval of their advisors (you need to consult with ALL faculty members whom you will ask to
        write letters, not just your primary advisor; and
        \item either have not participated in the job market process before,
    or participated but did not find a job (this is meant to reassure
    you, not to worry you; we have a very good record of students
    ultimately getting jobs).
    \end{enumerate}

    \phantomsection\hypertarget{DecideByWhen}{}
    
  \end{quote}

\item  When do I need to make a decision about whether I am going on
  the market?

  \begin{quote}
    If you are a 6th year student this year, you don't have a choice: If
    you have an acceptable job market paper in time, you must go on the
    market; if you don't have an acceptable job market paper in time, you
    are in deep trouble, since the department must approve your job market
    paper, you can't go on the market without a paper; and you can't stay
    in the program for a 7th year.  (See the graduate \href{https://www.econ2.jhu.edu/pdf/GraduateStudentHandbook_2023-2024.pdf}{handbook} for a fuller discussion of these issues.) 

    For other students (who would be in their 5th year, or very rarely 4th year):

    \begin{enumerate}
    \item You need to make a preliminary decision in late March or early April
      \begin{itemize}
      \item You need approval of your principal advisor
      \item You need to consult your other letter writers
      \end{itemize}
    \item You need to identify, and name to the JMPO
      \begin{itemize}
      \item Your principal advisor
      \item Two other advisors
      \end{itemize}
    \item If you are (now) a 4th year student, You should make a final decision by July 1.  
    \end{enumerate}
    
    \begin{comment}

      In the dept meeting on 2006-10-17 a question was raised about
      whether we are really serious about our policy, as stated by Larry,
      that ``there is no such thing as a 7th year student.''  

      The answer is yes.

      There was then some discussion about the relationship between going
      on the job market and remaining in the program, caused by the fact
      that there are currently three students (Tereanu, Svitil, Lu) who
      are clearly not ready for the job market but have not given up all
      hope of finishing a dissertation by August of next year.

      The resolution (subsequently clarified in emails on 2006-10-18 and
      2006-10-19 and an email to Zhou Lu on 2006-10-19 from CDC) was as
      follows.

      Our declared public policy (see the graduate \href{handbook}{http://www.econ2.jhu.edu/pdf/Econ_Grad_Handbook.pdf}) is that nobody can enroll for a 7th year
      in the program.  In practice, if in the opinion of an advisor and
      second advisor, a student's dissertation is essentially complete as
      of August of the 6th year, we may relax this rigor slightly in order
      to arrange a defense and final ``cleaning up'' of details - but \textit{  only} in the case where the dissertation is in all essentials
      complete.

      So, an important real and psychological point here is that failure
      to be included in the official list of candidates is almost (but not
      quite) equivalent to the end of all hope of obtaining a JHU PhD.  If
      a student uses wisely the time freed up by not participating in the
      job market to conduct an intense burst of high quality research, it
      is in principle quite possible that they will finish a dissertation
      despite their prior lack of progress.  The exclusion from the job
      market process may concentrate their minds in a way that has not
      happened before.

      Furthermore, if their dissertation defense occurs before the deadline for the `on-the-market' decision in the following year (mid October), the student can apply to jobs themselves (with the usual process, including advisors writing letters and giving them to the \JMStaff, etc), but without their name being officially included in list of `on-the-market' students.  They can also have a page on the dept website to which they can refer employers, but will not be included in the list of students ``on the market'' clickable from the main dept web page.  The situation will be explained candidly to employers as resulting from our inflexible rule that students must be on the market in their 6th year or earlier.  This is what we did with Farhan Hameed this year, and he did eventually get a couple of mediocre job offers, one of which he has accepted.

    \end{comment}


  \end{quote}
\item It's before July 1, and I'm not sure whether I'll be on the market or not. Which steps should I take (e.g.\ should I schedule a job seminar)?

  \begin{quote}
    Unless you are sure you are \textit{not} going to be on the market, you should behave as though you are sure you \textit{are}.  So, yes, go ahead and schedule a practice job talk, etc.

  \end{quote}
\item  Can I `partly' go on the job market?  That is, apply to a few
  selected employers (say, the World Bank's YP program) only, and 
  if that doesn't work go on the market more fully next year?

  \begin{quote}
    No.  This does not work.  The things you (and your advisors) need to do in order for you to
    get a job are basically the same: Get letters of recommendation, 
    write a respectable job market paper, etc.  If you do not do them
    properly, you won't get a job anywhere; and, if you do them well,
    you will have good options.  A half-baked application is worse than
    none at all, because you make a bad impression on that employer, and 
    employers talk to each other a lot, so that bad impression might damage
    your chances elsewhere in the future as well.  

    You can't go on the job market twice (with support from JHU, anyway; and, you won't
    get a job without support from JHU).

  \end{quote}
\item  Can I use a coauthored paper as my job market paper?

  \begin{quote}
    Yes - if the coauthor is another student, \textit{and that student is not
    also using the same paper as a job market paper}.  No, if the paper is with one of your
    advisors.  Maybe, if the paper is with someone who is not a student
    and not an advisor (and not famous).  You will need to discuss with your advisor.
    \hypertarget{Where-Should-I-Look-for-Jobs}{}

  \end{quote}

\item  Where should I look for jobs? 

  \begin{quote}
    There is one main source: \JOE.

    No other source that I know of is large enough to direct everyone to. (If any student finds a rich place that everyone should be looking at for job postings, please let me know).

    \hypertarget{Region-Specific-Job-Markets}{}
    
    There are some region-specific job markets that may be worth paying attention to if you are particularly interested in a job in the region in question.  For example, over the years there have been several efforts to produce a European job market patterned on the US market.  Starting in 2018 a new effort by the \href{https://www.eeassoc.org/}{European Economic Association} has been launched, which seems to have learned some lessons from (failed) past efforts to create a thick matching market for European institutions hiring economists. Some European institutions (universities; central banks) have said they have switched their recruiting efforts to the European from the American market.  In order to participate in the European job market you must get a EEA membership and register \href{https://www.europeanjobmarketofeconomists.org/}{here}.  The problem is that anyone can say in a Zoom interview they'd be delighted to have a job in Europe when a European institution interviews them, but then when they have offers from comparably ranked American institutions they tend to go to the American employer.  If you want to send a costly signal that you are really interested in jobs in Europe, you might consider flying to Europe for the EEA meetings (typically in late August), and try while there to contact institutions that have already advertised an available position.
 
    Similar logic probably applies to other regional or country-specifc markets.
  \end{quote}
  
\item What is the timing of posting of jobs on JOE? 

  \begin{quote}
    Each year is slightly different from the previous. See the table (credit: Daniel Garcia) 
    that shows the number of new jobs posted on JOE between different dates 
    during 2016's job cycle:
    \ifdvi\phantomsection\hypertarget{JobPostDates}{}\fi
    \begin{table}[h]
      \centering
      \begin{tabular}{|l|c|}
        \hline
        Dates                   & \# of job postings on JOE \\ \hline
        Aug 1 -- Oct 11   & 740                       \\ \hline
        Oct 11 -- Nov 1 & 340                       \\ \hline
        Nov 1 -- Dec 1 & 290                       \\ \hline
        Dec 1 -- Jan 1  & 115                       \\ \hline
      \end{tabular}
    \end{table}
  \end{quote}

\item How many references should I have?

  \begin{quote}
    Absolute minimum (and ideal number) of 3.  Absolute maxmimum of 5 (under
    extremely special circumstances).  

    You should list as references \textit{only} people who have \textit{agreed}
    to be references.  You will need three letter-writers, and they should know \textit{long in advance} that you will be asking them for a letter.  5 is the absolute
    maximum number of letters, and is NOT preferred - employers will not read 5 letters,
    except in rare cases.  3 is the optimal number.  4 is acceptable, and
    better than 3 if you really have 4 people who are very familiar with
    your work \textit{or} your qualifications, or if there is someone who can
    uniquely testify to a particular talent (brilliant work as a Fed RA,
    for example) that nobody else can see.  5 is reserved for cases where
    you have many people both inside the dept and outside who are
    thoroughly familiar with some aspect of your work (though it doesn't
    have to be your thesis; someone for whom you worked as an RA might
    write an excellent letter about how you performed).  You should \textit{not} have 5 references if some of those people are only vaguely
    acquainted with what you have done.  You should probably have at least
    two Hopkins faculty members.  (To be crystal clear, you can have three letter-writers and one or two other people who are listed as references in the sense that you provide their contact information in caes anyone wants to consult them.

    Remember that someone reviewing your file is not overcome with
    a thrill of anticipation when they see that you have 4 or 5 letters; they
    are overcome with a pang of annoyance at the extra work.  So don't foolishly
    imagine that ``more is always better;'' include the extra letters if
    there's a good reason, but not otherwise.


  \end{quote}
\item  Can I send different sets of letters to different employers?

  \begin{quote}
    If every student had different letters for every employer, it would be impossible for {\JMStaffName}~to get it all correct (and a huge amount of work even to try).

    {\JMStaffName}~is prepared to handle different sets of letters to different categories of employers (usually, academic vs nonacademic) but you need to explicitly label those categories for her, and make SURE your letter writers label the letters clearly in the different categories when they send them to her.  Like, maybe, letters for academic institutions should have \texttt{academic} in the name of the PDF.

    Depending on {\JMStaffName}'s bandwidth, you might ask for a \textit{few} special-case deviations from your usual set of letters.

    If the motivation is that you have a potential letter writer who has particular contacts at particular institutions that might be helpful, another (perhaps better) option is to have that writer email their contact at the institution directly, and send either a full letter or a personal email endorsement (or both).  This is more work for the letter writer, but is likely to be more effective than just having their letter included as (say) a fourth or, God forbid, fifth letter.

  \end{quote}
\item  How is the market for economists this year?

  \begin{quote}
    We never really know until the market is over.  But there is an enormous amount of idiosyncratic
    variation from subfield to subfield, person to person, and institution to
    institution, so the overall market conditions probably do not matter as much
    as you think, either for good or for ill.

  \end{quote}
\item  \ifthenelse{\boolean{MyNotes}}{\marginpar{\tiny }}{} 
  How many job applications should I send out?

  \begin{quote}
    First, a key principle: You must not apply to a job that you know you would not
    accept under any circumstances, even if it were your only offer.  If
    at the end of the job market process you have at least one job offer
    and you refuse it, you forfeit the right to any further help from us
    in finding another job subsequently.  %(Of course, we reserve the right to relax this rule and choose to help you if you can persuade us that your reasons for turning the job down are compelling).

    Given that, there is a range of opinion on how many applications to send.  Some students basically apply to every employer advertising a job for which they could conceivably qualify, which at a maximum might lead to as many as 100 applications.  Others cull the list.  We \textit{strongly encourage} you to apply to a minimum of 25 potential employers.  Note that every extra employer you apply to puts a burden not just on you but on the staff, who must send out multiple reference letters for each employer.  So please do not apply to zero-probability places.  It's a waste of everybody's time.

    We have found that under these guidelines, most students apply to around
    50-100 employers, though numbers can vary depending on circumstances 
    (e.g.\ a student with a Fulbright that requires them to apply only in 
    Europe may have a much more restricted set of potential choices,
    and therefore will apply to fewer employers than students who face
    no such restrictions).


    \ifdvi\phantomsection\hypertarget{FedHiringRules}{}\fi

  \end{quote}
\item  \ifthenelse{\boolean{MyNotes}}{\marginpar{\tiny }}{}  What are the Fed's rules about hiring persons who are not US citizens? 

  \begin{quote}
    For the Board of Governors, the answer is that it depends on the country you are from.  As of 2017, \href{https://www.federalreserve.gov/boarddocs/srletters/2006/SR0614a3.pdf}{the document linked here} which lists the eligible countries was found by Googling ``Does the Federal Reserve hire nonimmigrant aliens?'' 

    According to \href{https://www.federalreserve.gov/careers-faqs.htm}{this FAQ}, ``The Board will hire any qualified person who is a United States citizen or who is not a citizen but meets the requirements of the Immigration Reform and Control Act of 1986 (IRCA).  As required by law, the Board will not hire any person who is unable to satisfy the requirements identified in IRCA. In addition, citizens and nationals of the United States will receive preference in employment over equally qualified persons who are not citizens or nationals of the United States.''

    As I understand it, the Federal Reserve Banks can hire people from countries not in this list, but cannot give them ``Class 2'' clearance until they have served for two years, which means that the person cannot do much of what economists are hired at a Bank to do during that period.\footnote{For example, they cannot see the TealBook (the Fed's real-time assessment of economic conditions compiled before the FOMC meeting), and the preparation of which is a principal responsibility of economists at both the Board of Governors and at the Federal Reserve Banks.  It can \textit{never} give them ``Class 1'' clearance, so such employees are limited in what they can do (which makes them much less attractive to the Fed).}  This means that the Banks only rarely hire persons from countries not on the list, and generally only when that person has some unique skill or knowledge that cannot be obtained in any other way.

    As of the fall of 2017, it is my understanding that the Board and the Banks now sponsor people for Green cards after three years of employment.  (This seems to have changed over time, so if it is important to your job choice then you should verify it).

  \end{quote}
\item  \ifthenelse{\boolean{MyNotes}}{\marginpar{\tiny }}{} 
  When I am ready for my recommendation letters to be sent (AFTER I have
  sent my applications to employers), what is the process?

  \begin{quote}

    see
    
    {\pageurl/Recletters}

    for detailed answers

  \end{quote}
\item   Can there be more references than people writing the recommendation 
  letters? For example, people who are not writing a letter but are willing to provide 
  an evaluation of the applicant in case the potential employer contacts them?

  \begin{quote}
    The ``References'' part of your CV is basically a list of people from
    whom the employers will expect to receive a letter.  For references who
    are not Hopkins faculty members, include complete contact information
    (address, phone number, email address).

    If you want to indicate that there are other people (not in your
    ``references'' list) who could be contacted, including their contact
    information somewhere else on the CV is a very obvious suggestion to
    the reader that the person is happy to be contacted (e.g.\ if you
    worked as an RA for someone who would be happy to recommend you,
    don't just list their name when describing the job, list their email
    address and phone number).


  \end{quote}
\item  Do I need to write a different custom cover letter for every employer?
  
  \begin{quote}
    Yes - at least in the sense that you cannot write a generic ``Dear Potential Employer'' letter
    and send it to all 100 places you are applying.  Mail-merge software needs to be used to make sure
    that every letter has the name of the chair of the relevant search
    committee, the institution and department to which you are applying,
    and the nature of the job for which you are applying (assistant
    professor, economist, god-king for life, etc).  The \Templates~ directory contains templates that 
    you can use to create the letters from your \EMtt~list.  
    
    You should do more highly customized letters for those employers
    where a) you have a special connection; b) you have special
    interests or a particularly close match to the employer's need; or
    c) there is some other reason to think they might read the letter
    and find it persuasive.  

    In most cases, it may not be worth your time to do more than
    5-10 customized letters.  (And, the \textit{first sentence} of the letter should signal
    clearly that it is customized -- people reading applications will probably assume that the letter is \textit{not} customized and will only glance at it briefly; if the first sentence looks like boilerplate, they may miss whatever customizations you have incorporated later).

    \ifdvi\phantomsection\hypertarget{WhatIfEmployerContactsStudent}{}\fi
  \end{quote}
\item  What does it mean if an employer to whom I have not applied sends me an email asking me if I would be interested in applying?

  \begin{quote}
    This could mean any of several things.  One is that the employer has looked through the JHU candidates and has identified you as someone who they think might be an especially good fit (perhaps for reasons that are not obvious to you from their JOE posting, like the interests of existing faculty members).  Another is that they have heard something favorable about you from someone and want to get more information.  Yet another is that they might think that you ruled them out because you didn't know much about them but if you learned more about them you might find that they are better than you think.  (This is particularly true of schools that may have improved sharply in the recent past -- maybe as a result of an infusion of donor money, or hiring new faculty, or any of a host of other reasons, but whose improved quality has not yet been reflected in the (often badly out of date) rankings of departments on the internet).

    In any case, this is a signal of the school's interest in you, and generally it is a good idea to apply unless you have some strong reason not to do so (like, you would be ineligible to go there, or would not want to go there even if it were your only offer).  You can also seek your advisor(s) advice and that of the JMPO.


  \end{quote}

\hypertarget{record-and-post-your-job-talk}{}
\item  How does the video of the seminar work?
  
  \begin{quote}
    You should have identified a classmate (your `buddy') whose
    responsibility will be to help you during your seminar presentation.
    
    Before the seminar, you and your buddy should:
    \begin{enumerate}
    	\item (One week before) Create a zoom link and send it to the seminar organizer so that they can distribute it in the announcement for your presentation.
    	\item Share the link with the \href{\Notationpage}{JMCC} and \href{\Notationpage}{JMPO} so that they can add it to the JobMarket-KeyDates calendar. The JMPO might want to connect to your talk remotely so it is crucial for you to send him or her the link.
    	\item (Half an hour before) Arrive at the seminar room and set everything up. Join the zoom call. Test the cameras and microphones are transmitting to the zoom call. Verify that the clicker works. Load your presentation.
    	\item (5 minutes before) Start recording. You will ultimately need an mp4 file of your talk. You can record "locally," and zoom will produce this mp4 file for you.
    	\item Again, verify that everything works. Have your buddy join the zoom call and check the audio, video, and screen-sharing.
    \end{enumerate}
    
    Present your paper and end the recording when you are done.
    
    After the seminar, you have to upload the recording to a Panopto folder that the JMCC will create and provide a link for.
    \begin{enumerate}
    	\item Extract the recording from the seminar room computer using a USB stick. It is important for you to bring a USB stick to your presentation.
 		\item Using your own computer, upload your file to the Panopto folder:
 		\begin{enumerate}
 			\item Navigate to the Panopto folder using your web browser.
 			\item Click on the three "Create" button on the upper-left corner.
 			\item Click on "Upload media"
 			\item Add your file and make sure to select the "Econ-JobMarket-YYYY" folder provided by the JMCC.
 		\end{enumerate}
 		\item (If you want) Edit the recording to using Panopto to, for example, remove the dead time before and after the actual presentation in the recording.
 		\item Name your recording ``YourMoniker\_PracticeJobTalk''
    \end{enumerate}  

    Your `buddy' should also take detailed notes about the questions
    that come up during the seminar.  You should also ask all the other
    jobmarket candidates who attend your practice job seminar (and any
    friends you have who are not yet candidates) to take notes on how
    things go, and you should write up a summary of the combination of
    these notes (integrated with your own viewing of the video)
    afterwards.  This is the best way to be sure you absorb and process
    the feedback that you receive during the talk.  This is an
    exceptionally useful thing to do, and should probably take you at least
    a day; you should do it as soon as possible after the job talk, and
    make an appointment with your advisors the following week in which
    you share your writeup of the notes and sketch what you think you
    learned about how to improve.  Do NOT skimp on this; it is one of
    your best opportnities for making big improvements that you might
    otherwise never think of on your own.

  \end{quote}
\item  Should I apply to schools who list jobs in fields that are not exactly my field but are related to my
  field?  Like, suppose my thesis is on international trade and a school lists
  an opening in ``international relations.''

  \begin{quote}
    You will have to excercise judgment on this.  For example, in the case
    above, if the ``International Relations'' job is in a business
    school's economics department and the job description mentions
    international trade, then yes, it is reasonable to apply.  If it's in
    a public policy school's International Relations department and your
    sense that they are actually looking to hire someone to bloviate about
    ``globalization'' then it's probably a waste of time.  Your advisers
    should be able to give you good advice in specific cases.

  \end{quote}
\item  Some places ask for a statement of teaching and/or research
  interests.  Is this different than our 1 page
  dissertation summary?

  \begin{quote}
    A statement of teaching interests or teaching philosophy is
    definitely different from the dissertation summary.  There are some
    examples of previous students' statements in \Resources.  Many
    schools also ask for quantitative information about teaching
    evaluations.  We have provided a template for that too in
    \Templates.  Please provide feedback to the JMCC or JMPO if you
    think this template could be usefully modified -- it is fairly new.

    If the school asks for a statement of research interests it is probably
    a good idea to include such a statement separately from the dissertation
    abstract, though in some cases (use your judgement) the abstract will suffice.  Again, some examples are in \Resources.  
    The key thing you can communicate here is a sense of where your longer-term 
    interests are -- what do you anticipate working on next once your dissertation
    papers are published?  How does your work fit into broader developments in 
    the economic research literature?

    \ifdvi\phantomsection\hypertarget{PostOtherDocsQ}{}\fi
  \end{quote}
\item  The {\JOE} posting for many jobs asks for only a subset of all the materials I have assembled.  Should I upload everything anyway as ``other documents''?

  \begin{quote}
    No. The employers in question probably are trying to cut down on the massive amount of bandwidth/clutter they get from the process.  They have probably chosen what to ask for carefully.  There are only two kinds of content that you should upload even if not asked for: (1) your job market paper (they may have thought it was so obvious that they needed this that they did not explicitly list it); and (2) if you have any one thing that is especially notable that they might be particularly impressed or intrigued by (your sole-authored AER paper, say!) you should upload that.

  \end{quote}
\item  Will the department favor me over my classmate doing similar work?  Alternatively, how do I crush/avoid crushing my long-time friend in the job market?

  \begin{quote}
    This is a (much) less important issue than you might think, because even when schools say they are looking in ``all fields'' they usually are actually strongly biased toward filling some particular need.  Even if they say they are looking in ``health'' they may be really looking for an IO person in health, or really looking for a structural applied micro health person, or a public finance health person, or whatever.  Even at big institutions (the Fed; the IMF) the positions are available for particular groups doing particular things.  So in practice, there is very little real issue of ``student rivalry'' in the sense of worrying about whether other JHU students are your competitors for a given job, or whether the placement director might tilt the playing field somehow; in my (many) years as a placement director I have never seen a case where two JHU students were actually close substitutes for a given match.  As a result, my incentives are almost perfectly aligned with the students' incentives, in the sense that my job performance is measured by how well JHU students do on the market, and my incentive is to help find the right matches.

  \end{quote}

\item  It is fairly early (November or early December), and I have received several requests for interviews to my ``backup'' places but might want to cancel them if I have enough interviews at my ``preferred'' places later.  Is it OK to accept interviews at my ``backup'' places for now and then back out later if it turns out that I need to use the slots for better places?

  \begin{quote}
    Yes, you can back out later.  On the whole it is not wise to do so unless every moment of your possible interview time is filled and then someone calls who is preferred to some of your existing interviews.  Again, remember the principle that you are likely to be interacting with many of the same people throughout your career, and you don't want your first impression on them to be a bad one.  Also, you may ultimately only receive a job offer from your ``backup'' places, not the ones you hope for!

  \end{quote}
  \ifdvi\phantomsection\hypertarget{InterviewBestTime}{}\fi
  
\item  I've received an interview request for one of my most preferred employers. When is the best time to pick to do the interview?
  \begin{itemize}
  \item Morning, but not the first slot -- you and the interviewers will be awake (coffee will have kicked in) but not yet be tired
  \item The second day of not the first day of the conference
    \begin{itemize}
    \item You'll have had a few practice interviews to get in the groove, and will still be ``fresh'' and not robotic sounding
    \item The interviewers similarly will have had enough interviews to be in their groove, but will also still be fresh
    \end{itemize}
  \end{itemize}

\item  Should I ask whether my expenses will be reimbursed by schools that fly me out for an interview?
  \begin{quote}
    No.  Expenses are always reimbursed (if you keep good records and submit an expense report).  For this purpose, \url{http://expensify.com} is useful.

  \end{quote}

\item  What happens during a flyout?
  \begin{quote}
    Obviously, you will give your job market seminar.  But more is
    involved.  You generally meet individually with faculty at a
    fly-out.  Usually a few before your talk, then give your talk, then
    lunch, then meet with more faculty, then dinner.  Somewhere in there
    you may also meet with the Dean, so you might want to be prepared
    with a question or two for him/her (one could be something about
    their tenure policy).  Occasionally, at big schools, you might meet
    with groups of a few faculty (so that everyone has a chance to meet
    you).  The conversations are generally more relaxed and informal
    than the interviews at the meetings, and also more wide-ranging.
    This is where it really pays off to be an avid reader of the \textit{Economist},
    the \textit{Financial Times}, good economics blogs, and other news sources.
    If someone mentions that they read a recent story related to your 
    work in the \textit{Economist} and you haven't read it, you'll look like
    you're not really on top of your field.  If your conversation strays
    a long way from your dissertation but the faculty member is impressed 
    by your general knowledge, that's a very good thing.

    These meetings often also focus on where your research going from
    here, and maybe some discussion about teaching experience (especially
    at small liberal arts colleges).  It's impossible to prepare for
    everyone you're going to meet (especially since you usually don't get
    the schedule until late the evening before).  But in the meetings, you
    can also ask some of them what they are working on, how they like the
    school, what their experience has been with undergrad RAs, etc.  They
    also want to make sure that they would like you as a colleague.  (It is
    also a really long day, not that there's anything to do about that other
    than getting enough sleep the night before.)



  \end{quote}

\item  \ifthenelse{\boolean{MyNotes}}{\marginpar{\tiny }}{} 
  What happens if I don't get any job offers?
  \begin{quote}
    First, don't start worrying too early.  The way the job market works
    is as follows: The market decides who are the superstars for the year, and
    those people each get 25 offers.  The 22 places that have no chance of
    hiring them (along with the 3 who do have a chance) then wait for a
    month until they learn the candidate has turned them down.  They then
    make an offer to a person they think is another superstar, and wait for that person
    to select among their many offers.  Then, in March, they get
    realistic and start making offers to people they might actually be
    able to hire.  So if you don't have any job offers by Feb 15, don't
    worry.  (But do keep in touch with your advisors and with me during this
    period).

    Usually the candidates who have the most difficulty finding jobs are
    those with strong restrictions (usually geographical).  In other
    words, if you say ``I can only take a job in New Jersey'' then it can
    be very difficult - this is the only category of student who 
    have sometimes not been able to find a job in the last few years.

    As noted above, if things don't work out on the market this year, then
    you have the right to our help again next year.

  \end{quote}

\item  Does the department pay any of my expenses for the job market?

  \begin{quote}  
    At some schools, students going on the market are charged a fee to cover the costs of administrative assistance in the department's marketing of their students.  We provide the services of the JMPO, the JMCC, the job market staff person, the web liaison, and faculty who can help you in various other ways.  But that is the extent of our ability to help with the expenses associated with the market.  Think of your expenses as an investment!

    PS.\ Note that employers will pay for the travel of candidates they invite to give job talks. 

  \end{quote}

\item  How long should my job market paper be?

  \begin{quote}
    Maximum of 35 double-spaced pages (excluding figs, tables,
    appendices), min 12 point type, 1.25 inch margins.  If figs and
    tables are included in the text, maximum of 42 pages.   
    
    A wise person once said that a paper is not finished when nothing
    remains to be added; it is finished when nothing remains that could
    be taken away.
    
    If you think you can't fit everything into these guidelines, think
    again.  You've got to realize that \textit{nobody is going to read a
      huge fat paper just because you spent a lot of time writing it}.
    Get to the point.  No rambling pages of text designed to prove you've
    read every paper since the seminal contribution of Moses (1200 BC).

    There is one escape clause to the above: If necessary, you can have
    substantial online appendices containing tables, proofs, and other material, and
    refer to them briefly in the text.  This is an excellent way to
    establish points that may be necessary to the argument of your paper
    but not particularly interesting in themselves.   

  \end{quote}

\item  How long should the cover letter be?
  \begin{quote}
    One page.  Period.

  \end{quote}

\item  What should be in the abstract?
  \begin{quote}
    The \href{\bloburl/moniker-dir-make/Abstract.tex}{template document} has four elements:
    \begin{itemize}
    \item Abstract for the dissertation as a whole
    \item Abstract for job market paper
    \item Abstract for the next most interesting paper
    \item Abstract for your third paper
    \end{itemize}

    The purpose of the abstract is to paint a picture of yourself that will stick in the mind of the reader.  It is NOT to overawe the reader with your claims for the breathtaking advances you have made relative to the work of Einstein (2022), because you can't make a credible case for that in a few sentences (even if you really do prove that quantum theory and relativity are both wrong and you have solved all the problems of physics forever).

    It must be one page, but does not need to fill the page.  Min 12 point type, 1.25 inch margins.  Don't try to squeeze the maximum possible number of words onto the page.  It is \textbf{more important that it be readable} and look professional than that it mention everything you have done.

  \end{quote}

\item  What's to stop me from applying for a job in JOE even if my advisors don't think I'm ready?  I'll just send my application and hope that someone will hire me even without advisors' approval.  (Nobody has been bold enough to ask this question directly, but I'm certain that students have thought about it).

  \begin{quote}
    There is no point in applying to employers unless you have approval from advisors and a reasonable job market paper.  Every job posting has a large number of applications (like, 400).  The first thing every employer does is to look at the letters.  If there are no letters, or the letters basically say ``this candidate is not prepared,'' the employer will throw your application in the trash and look at the next applicant in the pile of 400 other people who applied for the job.  If your letters are adequate, the employer looks next at the job market paper.  If it is incoherent, the employer A) throws your application away, and B) mentally downgrades JHU for the purpose of future hiring.  Hence our absolute requirement that you have a coherent job market paper or you're not going on the market.

    Just to make it crystal clear, I will say it again: If your advisors
    are not prepared to write a letter that says you should be hired, no
    employer will hire you.  So there is no point in trying to apply on
    your own without your advisors' approval - it would be a waste of
    paper, money, and time on your part.  Your application will simply be
    ignored if it does not come with advisors' letters, and will be discarded
    if it comes with letters that say you are not ready.

    (Of course, I am referring here to jobs which are part of the usual PhD hiring process in economics.  There's no reason you can't search for another kind of job, say as a brain surgeon or an astronaut, which does not require a PhD in economics.  And of course you are welcome to apply to jobs that require only a masters degree in economics, for which you can reasonably request letters from anyone who has supervised you in the JHU program).

  \end{quote}

\item  After interviews, should I send a ``thank you'' email to the interviewers, or is that too pushy/annoying?
  \begin{quote}
    It is best NOT to send a generic email ``Dear x, Thanks for interviewing me.''  But, for your {\Stepspage/\#bestshot} schools it's OK, and maybe even a good idea, to send a BRIEF and SPECIFICALLY TAILORED message to your \href{\Stepspage/\#best-shot}{best shot} schools.  ``Professor Putin, Thank you for the suggestion that I look into the mortgage market of Burkina Faso because its institutional details provide a test of my theory.''

    \hypertarget{restricted-data}{}  
  \end{quote}

\item  My job market paper uses restricted data, and I must obtain approval from the provider of the data before posting it. But the deadline for posting job market materials has arrived. What should I do?

  \begin{quote}
    First point: You should submit the version of the paper you want approved LONG ENOUGH BEFOREHAND to have an approved version by the time we need to post the paper (early October).

    Options if you have not received approval by then:
    \begin{enumerate}
    \item Make a PDF document whose entire content is one page saying ``This paper uses restricted data and cannot be posted publicly until [whoever] has approved its release.''
    \item Make a PDF document that contains the first page of your JMP (including abstract) and the second page can read as in the prior item.
    \end{enumerate}

    The latter option (or a modified version that redacts any ``results'' from the abstract) is much better. If you make such abstract redactions, you should see whether you can get the provider of the data to quickly review either the redacted or the full (unredacted) abstract, faster than they may be able to review the whole paper.

  \end{quote}

\item I was a graduate Teaching Assistant. I would like to present evidence of my teaching effectiveness as a graduate TA but the Department has not been able to provide me with any type of teaching evaluation for this period. What should I do?

\begin{quote}
	You could think of running your own anonymous survey emailing it to the cohorts you TA'd for. \href{https://docs.google.com/forms/}{Google Forms} and the Deparment's mailing lists (i.e., \texttt{econ-Xthyear@jhu.edu}) are your friends. Be sure to make the survey ``fair:''
	\begin{enumerate}
		\item Obtain pre-approval from the course instructor.
		\item Make sure the survey is anonymous.
		\item Email the whole cohort, not subsets of particular students.
		\item When presenting aggregate statistics, report the response rate you obtained.
	\end{enumerate}
\end{quote}

\item After the pandemic (up to at least 2023-24) interviews have mostly shifted online. Is it still worth it to go to the ASSA meetings in January?

\begin{quote}
	The JMPO had this advice for 2023-24, which might change in the future.
	
	``
	Your time is at a premium in this period, revising your job market paper and making enough progress on your other papers so that if/when you do flyouts you will have accomplished as much as possible.  So, unless you have either already secured a job, or are on the meeting program, it is probably best to skip the ASSA meetings.
	''
\end{quote}

\item There is an annual visit to the IMF headquarters. What is the point of this visit and how does it work?

\begin{quote}
	
	In early September, the IMF reaches out to the JMPO asking about students that would be interested in applying to their Economist Program. They request the students' names, fields, and \textbf{CVs}. I is therefore important to be sure that, by early September, you have an updated CV and have made the decision of whether to apply to their Economist program.
	
	The actual visit usually happens in late October. During the visit:
	``There will be an IMF team comprised of the EP team, a senior Economist and a former EP to provide information about the program, work of the Fund and this year's recruitment process, allowing for a Q\&A session to ask us any questions they may have. The format will be a brief introduction of our team, followed by the presentation, and \textbf{preliminary interviews (for those that meet the EP eligibility criteria)}''---from this year's email.
	
	Their eligibility criteria stipulate that you must be be:
	\begin{itemize}
		\item A recent PhD graduate or be within one year of completing a PhD in a relevant field of study
		\item A national of one of our IMF member countries
		\item Under age 34 when joining the IMF EP program
	\end{itemize}
	
\end{quote}

\end{enumerate}

\end{document}
